---
title: "Setup Xenium Project Run 1- HPC"
author: "Drake Williams"
date: "`r BiocStyle::doc_date()`"
output: html_document
knit: (function(inputFile, encoding) { 
    rmarkdown::render(
        inputFile, encoding = encoding, 
        output_file = file.path(
            dirname(inputFile), paste0('01_Setup_HPC_',Sys.Date(),'.html'))) 
                })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = F,
                      message = F, 
                      out.width = "100%", 
                      fig.align='center')
options(width = 1200)
```

# Introduction

The objective of this notebook is to 
1. crop the regions of interest so that only one section is present in each region
2. remove unnecessary files to save space (all xeniumranger files and duplicates)
3. load xenium files into Seurat. I normally do this on the BRICS server but I am unable to transfer large files there now so instead I did this on Biowulf (HPC). 

## Load packages

```{r load packages}
library(pacman)
p_load(Seurat, data.table, scCustomize, sf,dplyr)
options(Seurat.object.assay.version = "v5")
options(future.globals.maxSize = 1e12)
source('/data/williamsdrw/xenium/xenHelpers.R')
```



## Crop regions from run 1, 0um
```{r crop region 1 5um, eval=FALSE}


##s1r1
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s1r1/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s1r1_a/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s1r1_a_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s1r1/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s1r1_b/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s1r1_b_coordinates.csv')

##s1r2
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s1r2/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s1r2/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s1r2_coordinates.csv')

##s1r3
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s1r3/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s1r3/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s1r3_coordinates.csv')

##s1r5
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s1r5/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s1r5/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s1r5_coordinates.csv')

##s1r6
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s1r6/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s1r6/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s1r6_coordinates.csv')

##s2r1
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r1/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r1_HV203/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r1_HV203_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r1/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r1_HV207/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r1_HV207_coordinates.csv')

##s2r2
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r2/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r2_HV137/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r2_HV137_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r2/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r2_HV184/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r2_HV184_coordinates.csv')

##s2r3
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r3/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r3/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r3_coordinates.csv')

##s2r4
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r4/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r4_a/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r4_a_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r4/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r4_b/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r4_b_coordinates.csv')

##s2r5
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r5/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r5/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r5_coordinates.csv')

##s2r7
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r7/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r7_a/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r7_a_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r7/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r7_b/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r7_b_coordinates.csv')

##s2r8
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r8/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r8_HV185/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r8_HV185_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r8/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r8_HV211/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r8_HV211_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r8/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r8_HV188_a/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r8_HV188_a_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r8/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r8_HV188_b/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r8_HV188_b_coordinates.csv')

##s2r9
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r9/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r9_a/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r9_a_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um/s2r9/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_0um_crop/s2r9_b/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_crop_coords/s2r9_b_coordinates.csv')

```

## Crop regions from run 1, 15um
```{r crop region 1 15um, eval=FALSE}


##s1r1
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s1r1',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s1r1_a/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s1r1_a_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s1r1/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s1r1_b/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s1r1_b_coordinates.csv')

##s1r2
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s1r2',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s1r2/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s1r2_coordinates.csv')

##s1r3
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s1r3',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s1r3/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s1r3_coordinates.csv')

##s1r5
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s1r5',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s1r5/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s1r5_coordinates.csv')

##s1r6
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s1r6',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s1r6/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s1r6_coordinates.csv')

##s2r1
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r1',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r1_HV203/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r1_HV203_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r1',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r1_HV207/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r1_HV207_coordinates.csv')

##s2r2
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r2',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r2_HV137/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r2_HV137_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r2',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r2_HV184/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r2_HV184_coordinates.csv')

##s2r3
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r3',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r3/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r3_coordinates.csv')

##s2r4
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r4',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r4_a/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r4_a_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r4',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r4_b/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r4_b_coordinates.csv')

##s2r5
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r5',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r5/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r5_coordinates.csv')

##s2r7
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r7',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r7_a/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r7_a_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r7',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r7_b/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r7_b_coordinates.csv')

##s2r8
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r8',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r8_HV185/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r8_HV185_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r8',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r8_HV211/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r8_HV211_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r8',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r8_HV188_a/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r8_HV188_a_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r8',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r8_HV188_b/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r8_HV188_b_coordinates.csv')

##s2r9
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r9',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r9_a/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r9_a_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um/s2r9',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_15um_crop/s2r9_b/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run1_5um/s2r9_b_coordinates.csv')

```



## Generate run1 seurat object list
```{r run1 seurat}
# get the full directory list to pick out folders with data to be analyzed
dirList <- list.dirs(path = paste0('/data/williamsdrw/xenium/run1_0um_crop'), full.names = TRUE, recursive = F)

region <- list.dirs(path = paste0('/data/williamsdrw/xenium/run1_0um_crop'), full.names = F, recursive = F)

print("These are the directories that will be used to generate a xenium seurat object. Double check for errors.")
print(dirList)

xenList <- list()
areaList <- data.frame()
for(i in seq_along(dirList)){
  obj <- LoadXenium(data.dir = paste0(dirList[i]), fov = region[i])
  obj$orig.ident <- region[i]
  all_files <- list.files(dirList[i], recursive = TRUE, full.names = TRUE)
  cells_file <- all_files[grepl("cells.csv.gz", all_files)]
  expanded_files <- list.files(paste0('/data/williamsdrw/xenium/run1_15um_crop/',region[i]), recursive = TRUE, full.names = TRUE)
  expanded_cellfile <- expanded_files[grepl("cells.csv.gz", expanded_files)]
  expanded_kmeans <- expanded_files[grepl("clusters.csv", expanded_files)]
  expanded_de <- expanded_files[grepl("expression.csv", expanded_files)]
  areaMeta <- obtain_cell_area(cells_file)
  expanded_area <- kmeans_area(expanded_cellfile,
                               expanded_kmeans,
                               expanded_de)
  expanded_area$orig.ident <- region[i]
  obj <- AddMetaData(obj, areaMeta)
  obj <- subset(obj, nCount_Xenium > 0)
  xenList[i] <- obj
  areaList <- bind_rows(areaList, expanded_area)
}

```

## Crop regions from run 2, 0um
```{r crop region 2 5um, eval=FALSE}

##HV137
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV137A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV137A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV137A_coordinates.csv')

##HV140
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV140A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV140A_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV140A_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV140A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV140A_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV140A_B_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV140A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV140A_C/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV140A_C_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV140A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV140A_D/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV140A_D_coordinates.csv')

##HV160
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV160C/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV160C_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV160C_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV160C/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV160C_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV160C_B_coordinates.csv')

##HV171
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV171A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV171A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV171A_coordinates.csv')

##HV192
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV192A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV192A_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV192A_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV192A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV192A_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV192A_B_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV192A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV192A_C/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV192A_C_coordinates.csv')

##HV198B
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV198B/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV198B_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV198B_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV198B/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV198B_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV198B_B_coordinates.csv')

##HV191A
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV205B_HV191A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV191A_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV191A_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV205B_HV191A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV191A_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV191A_B_coordinates.csv')

##HV205B
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV205B_HV191A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV205B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV205B_coordinates.csv')

##HV205A
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/LAD11_HV205A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV205A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV205A_coordinates.csv')

##HV219AB
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV219AB/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV219AB_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV219AB_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV219AB/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV219AB_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV219AB_B_coordinates.csv')

##HV219AS
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV219AS/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV219AS_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV219AS_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV219AS/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV219AS_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV219AS_B_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV219AS/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV219AS_C/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV219AS_C_coordinates.csv')

##HV220
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV220A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV220A_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV220A_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/HV220A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/HV220A_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/HV220A_B_coordinates.csv')

##LAD1
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/LAD1/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/LAD1/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD1_coordinates.csv')

##LAD7
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/LAD7A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/LAD7A_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD7A_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/LAD7A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/LAD7A_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD7A_B_coordinates.csv')

##LAD8
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/LAD8/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/LAD8/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD8_coordinates.csv')

##LAD11
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/LAD11_HV205A/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/LAD11/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD11_coordinates.csv')

##LAD132B
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um/LAD132B/outs/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_0um_crop/LAD132B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD132B_coordinates.csv')

```
## Crop regions from run 2, 15um
```{r crop region 2 15um, eval=FALSE}

##HV137
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV137A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV137A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV137A_coordinates.csv')

##HV140
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV140A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV140A_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV140A_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV140A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV140A_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV140A_B_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV140A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV140A_C/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV140A_C_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV140A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV140A_D/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV140A_D_coordinates.csv')

##HV160
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV160C/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV160C_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV160C_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV160C/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV160C_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV160C_B_coordinates.csv')

##HV171
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV171A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV171A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV171A_coordinates.csv')

##HV192
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV192A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV192A_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV192A_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV192A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV192A_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV192A_B_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV192A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV192A_C/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV192A_C_coordinates.csv')

##HV198B
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV198B/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV198B_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV198B_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV198B/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV198B_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV198B_B_coordinates.csv')

##HV191A
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV205B_HV191A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV191A_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV191A_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV205B_HV191A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV191A_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV191A_B_coordinates.csv')

##HV205B
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV205B_HV191A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV205B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV205B_coordinates.csv')

##HV205A
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/LAD11_HV205A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV205A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV205A_coordinates.csv')

##HV219AB
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV219AB/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV219AB_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV219AB_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV219AB/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV219AB_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV219AB_B_coordinates.csv')

##HV219AS
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV219AS/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV219AS_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV219AS_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV219AS/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV219AS_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV219AS_B_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV219AS/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV219AS_C/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV219AS_C_coordinates.csv')

##HV220
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV220A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV220A_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV220A_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/HV220A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/HV220A_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/HV220A_B_coordinates.csv')

##LAD1
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/LAD1/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/LAD1/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD1_coordinates.csv')

##LAD7
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/LAD7A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/LAD7A_A/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_5um/LAD7A_A_coordinates.csv')
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/LAD7A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/LAD7A_B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD7A_B_coordinates.csv')

##LAD8
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/LAD8/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/LAD8/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD8_coordinates.csv')

##LAD11
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/LAD11_HV205A/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/LAD11/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD11_coordinates.csv')

##LAD132B
cropRegion_complex('/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um/LAD132B/outs',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_15um_crop/LAD132B/',
           '/gpfs/gsfs12/users/williamsdrw/xenium/run2_crop_coords/LAD132B_coordinates.csv')

```

## Generate run2 seurat object list, excluding LAD
```{r run2 seurat}
# get the full directory list to pick out folders with data to be analyzed
dirList <- list.dirs(path = paste0('/data/williamsdrw/xenium/run2_0um_crop'), full.names = TRUE, recursive = F)
dirListHP <- dirList[!grepl("LAD", dirList)]

region <- list.dirs(path = paste0('/data/williamsdrw/xenium/run2_0um_crop'), full.names = F, recursive = F)
regionHP <- region[!grepl("LAD", region)]


print("These are the directories that will be used to generate a xenium seurat object. Double check for errors.")
print(dirListHP)

xenList2 <- list()
for(i in seq_along(dirListHP)){
  obj <- LoadXenium(data.dir = paste0(dirListHP[i]), fov = regionHP[i])
  obj$orig.ident <- regionHP[i]
  all_files <- list.files(dirListHP[i], recursive = TRUE, full.names = TRUE)
  cells_file <- all_files[grepl("cells.csv.gz", all_files)]
  expanded_files <- list.files(paste0('/data/williamsdrw/xenium/run2_15um_crop/',regionHP[i]), recursive = TRUE, full.names = TRUE)
  expanded_cellfile <- expanded_files[grepl("cells.csv.gz", expanded_files)]
  expanded_kmeans <- expanded_files[grepl("clusters.csv", expanded_files)]
  expanded_de <- expanded_files[grepl("expression.csv", expanded_files)]
  areaMeta <- obtain_cell_area(cells_file)
  if(!identical(expanded_files, character(0))){
   expanded_area <- kmeans_area(expanded_cellfile,
                               expanded_kmeans,
                               expanded_de)
   expanded_area$orig.ident <- regionHP[i] 
   areaList <- bind_rows(areaList, expanded_area)
  }
  obj <- AddMetaData(obj, areaMeta)
  obj <- subset(obj, nCount_Xenium > 0)
  xenList2[i] <- obj
}

combined <- Merge_Seurat_List(c(xenList,xenList2), add.cell.ids = seq(1,length(c(xenList2,xenList)),1), merge.data = F)
combined <- Store_Misc_Info_Seurat(seurat_object = combined, data_to_store = areaList, data_name = "expanded_area_kmeans2")

fwrite(areaList, file='/data/williamsdrw/xenium/expandedArea_cellLevel_metadata_HP.csv')
saveRDS(combined, file=paste0("/gpfs/gsfs12/users/williamsdrw/xenium/", Sys.Date, "-run1_run2_0um_HP.rds"))
```

```{r}
dirListLAD <- dirList[grepl("LAD", dirList)]
regionLAD <- region[grepl("LAD", region)]

print("These are the directories that will be used to generate a xenium seurat object. Double check for errors.")
print(dirListLAD)

areaList2 <- list()
xenList3 <- list()
for(i in seq_along(dirListLAD)){
  obj <- LoadXenium(data.dir = paste0(dirListLAD[i]), fov = regionLAD[i])
  obj$orig.ident <- regionLAD[i]
  all_files <- list.files(dirListLAD[i], recursive = TRUE, full.names = TRUE)
  cells_file <- all_files[grepl("cells.csv.gz", all_files)]
  expanded_files <- list.files(paste0('/data/williamsdrw/xenium/run2_15um_crop/',regionLAD[i]), recursive = TRUE, full.names = TRUE)
  expanded_cellfile <- expanded_files[grepl("cells.csv.gz", expanded_files)]
  expanded_kmeans <- expanded_files[grepl("clusters.csv", expanded_files)]
  expanded_de <- expanded_files[grepl("expression.csv", expanded_files)]
  areaMeta <- obtain_cell_area(cells_file)
  if(!identical(expanded_files, character(0))){
   expanded_area <- kmeans_area(expanded_cellfile,
                               expanded_kmeans,
                               expanded_de)
   expanded_area$orig.ident <- regionLAD[i] 
   areaList2 <- bind_rows(areaList2, expanded_area)
  }
  obj <- AddMetaData(obj, areaMeta)
  obj <- subset(obj, nCount_Xenium > 0)
  xenList3[i] <- obj
}

combined2 <- Merge_Seurat_List(c(xenList3), add.cell.ids = seq(1,length(c(xenList3)),1), merge.data = F)
combined2 <- Store_Misc_Info_Seurat(seurat_object = combined2, data_to_store = areaList2, data_name = "expanded_area_kmeans2")

fwrite(areaList2, file='/data/williamsdrw/xenium/expandedArea_cellLevel_metadata_LAD.csv')
saveRDS(combined2, file=paste0("/gpfs/gsfs12/users/williamsdrw/xenium/", Sys.Date(), "-run1_run2_0um_LAD.rds"))

```


## Session Info
```{r session info}
sessionInfo()
```

